---
title: "Erin The Fourth of Tampa"
author: "Juliet Nuccio"
format: html
editor: visual
---

## Load Packages

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
```

## Prompt

The year is 1973. UC Berkeley has been sued for sex-based admissions \# discrimination. Specifically, the plaintiff claims that women are discrimina-# ted against in admissions policies. The court has provided admissions data to \# you and tasked you with answering the question -- if a student applies to the \# school, will their gender play a significant role in their chance of admission?# (We are keeping the analysis limited to this dataset)

## load the data

```{r}
url <- "https://waf.cs.illinois.edu/discovery/berkeley.csv"
data <- read.csv(url)
```

## EDA

```{r}
head(data)
summary(data)
```

## Calculate the admission rates by gender

```{r}
table(data$Gender, data$Admission)

admission_rates_gender <- data %>%
  group_by(Gender) %>%
  summarize(
    Total = n(),
    Accepted = sum(Admission == "Accepted"),
    Admission_Rate = Accepted / Total
  )
ggplot(admission_rates_gender, aes(x = Gender, y = Admission_Rate)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Admission Rates by Gender", 
       x = "Gender", 
       y = "Admission Rate") +
  theme_minimal()
```

## Admissions rate by major

```{r}
admission_rate_major <- data %>%
  select(Major, Admission) %>%
  group_by(Major) %>%
  summarize(
    Total = n(),
    Accepted = sum(Admission == "Accepted"),
    Admission_Rate = Accepted / Total
  )
ggplot(admission_rate_major, aes(x = Major, y = Admission_Rate)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Admission Rates by Major", 
       x = "Major", 
       y = "Admission Rate") +
  theme_minimal()
```

## Applications by Gender, Major

```{r}
applications_gender_major <- data %>%
  select(Major, Gender) %>%
  group_by(Major, Gender) %>%
  summarize(
    Total = n()
  )
ggplot(applications_gender_major, aes(x = Major, y = Total, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Total Applications by Major and Gender", 
       x = "Major", 
       y = "Applications") +
  theme_minimal()
```

## See admission rates broken down by gender and major

```{r}
admission_rates_gender_major <- data %>%
  group_by(Major, Gender) %>%
  summarize(
    Total = n(),
    Accepted = sum(Admission == "Accepted"),
    Admission_Rate = Accepted / Total
  )

head(admission_rates_gender_major)

ggplot(admission_rates_gender_major, aes(x = Major, y = Admission_Rate, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Admission Rates by Major and Gender", 
       x = "Major", 
       y = "Admission Rate") +
  theme_minimal()
```

## See the admissions rate gender difference by major

```{r}
admission_rate_diff <- admission_rates_gender_major %>%
  select(Major, Gender, Admission_Rate) %>%
  pivot_wider(names_from = Gender, values_from = Admission_Rate) %>%
  mutate(Difference = M - F)

ggplot(admission_rate_diff, aes(x = Major, y = Difference, fill = Difference > 0)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("dark green", "skyblue"), 
                    labels = c("Female Higher", "Male Higher")) +  # Red for below 0, skyblue for above 0
  labs(title = "Difference Between Male and Female Admission Rates by Major",
       x = "Major", 
       y = "Difference (Male - Female)") +
  geom_hline(yintercept = 0, linetype = "dotted", color = "black", size = 1) +  # Add horizontal dotted line at y = 0
  theme_minimal() +
  theme(legend.title = element_blank())
```
